---
title: "Linear mixed models & post hoc analyses"
output: html_notebook
---
# Load packages up

```{r}
library(lme4)      # For mixed-effects models
library(lmerTest)  # For p-values with lmer()
library(tidyverse) # For data manipulation

```

#Load my data
#I just want session 1 as my analyses foccus around that

```{r}
data <- read_delim("~/Documents/master_internship/outputs_final_data/final_df.csv", delim = ";")
data1 <- filter(data, session == "ses-mri01")
colnames(data1)
head(data1)
```
```{r}
data1 <- data1 %>%
  mutate(
    stimuli_type = factor(stimuli_type, levels = c("Angry",    "Happy" )),
    response      = factor(response,      levels = c("Approach","Avoid" )),
    volatility    = factor(volatility,    levels = c("stable",  "volatile"))
  )
```
# Drop rows with true NAs in any model variable
# (response has 114 NAs we’ll remove; error_rate should have none)
```{r}
data_mod <- data1 %>%
  select(subject, error_rate, response, volatility, stimuli_type) %>%
  filter(
    !is.na(error_rate),
    !is.na(response),
    !is.na(volatility),
    !is.na(stimuli_type)
  ) %>%
  droplevels()

```

# Quick sanity checks

```{r}
nrow(data_mod)   # should be 7260 − 114 = 7146
lapply(data_mod[c("response","volatility","stimuli_type")], levels)
```
#1: Model 1: the effect of my dependent variables on error_rate & subject 

```{r}
m_max <- lmer(
  error_rate ~ response
             * volatility
             * stimuli_type
    + (1
       + response
       + volatility
       + stimuli_type
       || subject),
  data    = data_mod,
  control = lmerControl(
    optimizer   = "bobyqa",
    calc.derivs  = FALSE,
    optCtrl      = list(maxfun = 2e5)
  ),
  REML    = FALSE
)
```
#let's just see the summary of this stats

```{r}
summary(m_max)
```
```{r}
rePCA(m_max)
```
# 1. Fit a model without the stimuli_type random slopes:

```{r}
m_simp1 <- lmer(
  error_rate ~ response
             * volatility
             * stimuli_type
    + (1
       + response
       + volatility
       || subject),      # dropped stimuli_type slopes
  data    = data_mod,
  control = lmerControl(
    optimizer   = "bobyqa",
    calc.derivs  = FALSE,
    optCtrl      = list(maxfun = 2e5)
  ),
  REML    = FALSE
)

```

#Let's compare the two models to see which would be a better fit

```{r}
anova(m_max, m_simp1, refit = FALSE)
```
#p > .20 as  p = .884
#From this it appears that I don't loose much by dropping stimuli_type as a slope this I shall choose m_simp1 as my model and run posthoc tests 

#Now going to SOP
# I will first test out Volatility and error rate: Lower error rates fo stable vs volatile trials

```{r}
library(emmeans) 

```
# To observe the average effect of volatility accross all stimuli types

```{r}
emm_vol    <- emmeans(m_simp1, ~ volatility)
pairs(emm_vol, adjust = "none")

```
#It appears that the error rate is 0.0665 times higher accuracy in the stable compared to the volatile conditions. Also is significant. As z=5.870 the observed effect is highly unlikely under the null hypothesis. 


#Here we are looking at interactions 
```{r}
emm_vol_by_stim <- emmeans(m_simp1, ~ volatility | stimuli_type)
contrast(emm_vol_by_stim, method="pairwise", adjust="none")
```
##Now I wish to look at Congruency + error rate: lower error rates for congruent vs incongruent trials
#congruency is the interaction between the response and the stimuli type 

```{r}
emm_rt  <- emmeans(m_simp1, ~ response * stimuli_type)
```
# Now I need to define a congruency contrast this are my levels: 
#1. response=Approach, stimuli_type=Angry
#2. Approach, Happy
#3. Avoid, Angry
#4. Avoid, Happy
#Let’s say “congruent” = (Approach–Happy + Avoid–Angry) / 2
#“incongruent” = (Approach–Angry + Avoid–Happy) / 2
# so here I am doing a mean 

```{r}
cong_contrast <- list(
  congruency = c(-1, +1, +1, -1)/2
)

```

#Now I apply it:

```{r}
contrast(emm_rt, method = cong_contrast, adjust = "none")
```
#according to this congruent conditions see pretty much no difference in accuracy to incongruent ones. 
#Because of this I will test the effect of the contrast on volatility also as this differs from my expectation by looking at my plotted raw data
#3.1) Volatility + congruency and error rate: higher error rate differences for congruency in the stable compared to the volatile condition 

```{r}
library(emmeans)

# Get the 4 cell means of response×stimuli_type *within* each volatility
emm_rt_by_vol <- emmeans(m_simp1, ~ response * stimuli_type | volatility)

#Define the same congruency contrast
cong_contrast <- list(congruency = c(-1, +1, +1, -1)/2)

contrast(emm_rt_by_vol, method = cong_contrast, adjust = "none")

```
#Indeed accuracy is lower in the stable condition compared to the volatile condition, thus error rates are higher in the stable condition compared to the volatile one. However in the stable condition the p value is >0.05 thus non significant. 

##Now I wish to test my hypothesis on response time instead of error_rate(accuracy) 

#For this I am using the same model as before but substituting my independent variables:
```{r}
data_rt <- data1 %>%
  select(subject, RT_s, response, volatility, stimuli_type) %>%
  filter(
    !is.na(RT_s),
    !is.na(response),
    !is.na(volatility),
    !is.na(stimuli_type)
  ) %>%
  droplevels()
```


```{r}
m_rt <- lmer(
  RT_s ~ response
             * volatility
             * stimuli_type
    + (1
       + response
       + volatility
       || subject),      # dropped stimuli_type slopes
  data    = data_rt,
  control = lmerControl(
    optimizer   = "bobyqa",
    calc.derivs  = FALSE,
    optCtrl      = list(maxfun = 2e5)
  ),
  REML    = FALSE
)
```
#Singularity check
```{r}
summary(m_rt)
```
# I don't see issues here so I shall proceed with my postdoc analyses

#Congruency + RT: Faster response time for congruent vs incongruent trials
```{r}
emm_rt  <- emmeans(m_rt, ~ response * stimuli_type)
```
# I shall now apply the samne contrast as before for congruency

```{r}
cong_contrast <- list(
  congruency = c(-1, +1, +1, -1)/2
)
```

```{r}
contrast(emm_rt, method = cong_contrast, adjust = "none")
```
#I see a small effect of congruency on reaction time where RT_s is a bit slower for incongruent trials and this effect is significant. 

#Volatility + congruency and RT: Higher differences in RT for volatility in the congruent compared to the incongruent conditions
 
```{r}
emm_rt_by_vol <- emmeans(m_rt, ~ response * stimuli_type | volatility)
contrast(emm_rt_by_vol, method = cong_contrast, adjust = "none")

```
# I see faster reaction times for both congruency in the stable and volatile condition compare dto incongruent, yet fastest reaction times for congruency in the volatile condition with these effects being statisstically significant 
