---
title: "Linear mixed models & post hoc analyses"
output: html_notebook
---
# Load packages up

```{r}
library(lme4)      # For mixed-effects models
library(lmerTest)  # For p-values with lmer()
library(tidyverse) # For data manipulation

```

#Load my data
#I just want session 1 as my analyses foccus around that

```{r}
data <- read_delim("~/Documents/master_internship/outputs_final_data/final_df.csv", delim = ";")
data1 <- filter(data, session == "ses-mri01")
colnames(data1)
head(data1)
```
```{r}
data1 <- data1 %>%
  mutate(
    stimuli_type = factor(stimuli_type, levels = c("Angry",    "Happy" )),
    response      = factor(response,      levels = c("Approach","Avoid" )),
    volatility    = factor(volatility,    levels = c("stable",  "volatile"))
  )
```
# Drop rows with true NAs in any model variable
# (response has 114 NAs we’ll remove; error_rate should have none)
```{r}
data_mod <- data1 %>%
  select(subject, error_rate, response, volatility, stimuli_type) %>%
  filter(
    !is.na(error_rate),
    !is.na(response),
    !is.na(volatility),
    !is.na(stimuli_type)
  ) %>%
  droplevels()

```

# Quick sanity checks

```{r}
nrow(data_mod)   # should be 7260 − 114 = 7146
lapply(data_mod[c("response","volatility","stimuli_type")], levels)
```
#1: Model 1: the effect of my dependent variables on error_rate & subject 

```{r}
m_max <- lmer(
  error_rate ~ response
             * volatility
             * stimuli_type
    + (1
       + response
       + volatility
       + stimuli_type
       || subject),
  data    = data_mod,
  control = lmerControl(
    optimizer   = "bobyqa",
    calc.derivs  = FALSE,
    optCtrl      = list(maxfun = 2e5)
  ),
  REML    = FALSE
)
```
#let's just see the summary of this stats

```{r}
summary(m_max)
```
```{r}
rePCA(m_max)
```
# 1. Fit a model without the stimuli_type random slopes:

```{r}
m_simp1 <- lmer(
  error_rate ~ response
             * volatility
             * stimuli_type
    + (1
       + response
       + volatility
       || subject),      # dropped stimuli_type slopes
  data    = data_mod,
  control = lmerControl(
    optimizer   = "bobyqa",
    calc.derivs  = FALSE,
    optCtrl      = list(maxfun = 2e5)
  ),
  REML    = FALSE
)

```

#Let's compare the two models to see which would be a better fir

```{r}
anova(m_max, m_simp1, refit = FALSE)
```
#p > .20 as  p = .9268
#From this it appears that I don't loose much by dropping stimuli_type as a slope this I shall choose m_simp1 as my model and run posthoc tests 

#Now going to SOP
# I will first test out Volatility and error rate: Lower error rates fo stable vs volatile trials

```{r}
library(emmeans)
```




